---
title: "Hate Crime in Toronto"
subtitle: "Deep Investigation in Hate Crime"
author: 
  - Jingtao Luo
thanks: "Code and data are available at: https://github.com/JerryLuoJT/Hate_Crime_Toronto"
date: today
date-format: "DD MMMM YYYY"
abstract: "This paper analyzed the Hate Crime Open Data, which is free and available to the public on Open Data Toronto. The paper investigated distribution of hate crime in Toronto based on year, reason and neighbourhood. In summary, the hate crime has increased in last 6 years, and the most often reason is religion bias, and the Yonge-Bay Corridor has the largest amount of hate crimes. These findings can help the government understand hate crime better, and thus protect the citizens.  "
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# install.packages("tidyverse")
# install.packages("ggplot2")
# install.packages("dplyr")
# install.packages("readr")
# install.packages("janitor")
# install.packages("opendatatoronto")
# install.packages("knitr")
# install.packages("here")
library(tidyverse)
library(ggplot2)
library(dplyr)
library(readr)
library(janitor)
library(opendatatoronto)
library(knitr)
library(here)

#read the cleaned hate crime data
data <- read_csv(here::here("data/analysis_data/cleanned_data.csv"))
#generate .gitignore
install.packages("usethis")
usethis::git_vaccinate()


```


# Introduction

Toronto, often celebrated as one of the most multicultural cities in the world, stands as a beacon of diversity and inclusion. With over half of its population identifying as a visible minority and more than 180 languages spoken, the city reflects a rich tapestry of cultural, ethnic, and religious backgrounds. It has long been regarded as a model for successful urban integration, where immigrants and marginalized communities contribute to the city’s vibrant social, economic, and cultural fabric. However, beneath this inclusive narrative, Toronto also faces challenges in addressing the rising threat of hate crimes.

According to government of Canada [@govCan], hate crime is defined as criminal acts done by a person who is motivated by an extreme bias or hatred towards a particular social group. Research studies show that hate crimes cause “disproportionate harm” to individual victims as well as other members of the community belonging to the targeted social group. These crimes lead to not only physical but also psycho-emotional hurt. For example, if the assault occurs because the victim's race, then victim may live in fear of being victimized again for the rest of his life because a person's race can never be changed. 

Despite Toronto’s inclusive policies and progressive reputation, the city has not been immune to the global rise in hate-fueled violence. In recent years, incidents targeting specific communities—including Muslim, Jewish, Black, and LGBTQ+ populations—have underscored the persistent challenges that marginalized groups face. While hate crimes constitute only a fraction of overall crime, their impacts are disproportionately harmful, not only to direct victims but also to the broader communities they belong to, fostering fear, division, and alienation.

This paper aims to analyze the hate crimes in the following perspectives: when, where and why. When means the amount of crimes happened in each year. Where means the neighborhood of the hate crime occurred. And why is the reason that triggered this hate crime.

The Data @sec-data of this paper explores the `opendatatoronto` data, with tools in R and methods used to analyze the data and some observations from the new dataset. The Results @sec-results introduces more observations found from the data analysis by visualizing the data. The Discussion @sec-disc ties back the findings with the situation in real world. Lastly, the Conclusion @sec-concl summarizes key findings and gives insight for further study.

# Data {#sec-data}

All relevant data can be found in the Open Data Toronto portal, and extracted using the `opendatatoronto` library for `R` [@rOpendatatoronto]. In particular, the original dataset includes the information about each hate crime, investigated by the Hate Crime Unit by reported date since 2018, and its latest update was on 17 September, 2024. The information incluedes each crime's offence number, the exact date of occurance, the exact date of report, the police division where offence happened, the reason caused this hate crime, the neighborhood of such crime occurred, and if there is people got arrested.

## Data Tools

The data was extracted, cleaned and analysed using `R` [@citeR], with functions from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `opendatatoronto` [@rOpendatatoronto], `knitr` [@rKnitr], and `here` [@rHere].

## Overview of Dataset

The selected dataset has selected some of the variables that will be discussed later in the paper. The selected variables are id, year of crime occurred; reasons of hate crime: age, mental or physical disability, race, ethnicity, language, religion, sexual orientation, gender; and the neighborhood of this crime occurred. See `Table 1` for an overview of the data.

```{r}
#| label: hc_overview
#| fig-cap: 'Table 1: Sample of Hate Crime Open Data'
#| echo: false
# use 'knitr' generate a table of cleaned data for overview
knitr::kable(head(data[, 1:4]), "simple",   col.names = c('ID', 'Year', 'Bias Type','Neighbourhood'))

```

The column "Bias Type" is not on the original dataset, instead, the original one has variables for each type of bias, I use the "Bias Type" variable which stores the name of bias(data type: character) so this is easier for later calculation. In the original dataset, there are one more bias called multiple bias, which indicates that the crime is caused by more than one bias. Multiple bias may cause over-counting in other bias, so for easier calculation I didn't select crimes with multiple bias. Also, in the Neighbourhood variable, I find that some of the crimes didn't have information about neighbourhood, so I removed crime with "NSA" in the 'neighbouhood_158' column.

## Distribution of Hate Crimes per Year
The average number of hate crimes per year is 193. However, the range of crime is very wide: from 111 in 2018 up to 329 in 2023.
```{r}
#| label: hc-by-year
#| fig-cap: "Table 2: Count of Hate Crime by Year"
#| echo: false
#| 
# count each row by year, and store the value in a new dataset
hc_by_year <- data %>% group_by(occurrence_year) %>%
  summarize(count = n())
# use 'knitr' generate how many crime per year
knitr::kable(hc_by_year,
             col.names = c('Year', 'Total Crimes'))

```

## Distribution of Hate Crimes by Bias
Here we want to count the number of hate crime based on the bias that caused it. In the raw dataset, as I have mentioned before, we have 8 bias. However, after we delete the crimes with multiple bias or don't have the neighborhood information, we only have 6 bias in the dataset: age bias and disability bias are not in the cleaned dataset anymore.
```{r}
#| label: hc-by-bias
#| fig-cap: "Table 3: Count of Hate Crime by Bias"
#| echo: false
#| 
# count each row by bias type, and store the value in a new dataset
hc_by_bias <- data %>% group_by(bias_type) %>%
  summarize(count = n())
# use 'knitr' generate how many crime by bias types
knitr::kable(hc_by_bias,
             col.names = c('Bias Type', 'Total Crimes'))


```
The average number of crimes for each bias is 193.5. The range of number of crimes in different bias vary largely, from 3 to 516, this make me consider whether we have outlier. The lower bound and upper bound I calculated(See 'Table 4') indicating no outlier in this dataset.
```{r}
#| label: hc-by-bias-IQR
#| fig-cap: "Table 4: Bounds of Table 3's Data "
#| echo: false

#Check if there is outlier
#Calculate Q1, Q3, and IQR
Q1 <- as.numeric(quantile(hc_by_bias$count, 0.25)) 
Q3 <- as.numeric(quantile(hc_by_bias$count, 0.75))
IQR <- Q3 - Q1

#Find bounds
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR
#Summarize the data into a data frame
summary_table <- data.frame(summary_table <- data.frame(
  Q1 = Q1,
  Q3 = Q3,
  IQR = IQR,
  Lower_Bound = lower_bound,
  Upper_Bound = upper_bound
)
)

# Display the table using knitr::kable
knitr::kable(summary_table)

```

## Distribution of Hate Crime by Neighbourhood
I am curious about which neighbourhood in Toronto has the most hate crimes, so I count the data by variable 'neighbourhood_158', which store the neighbourhood name of each crime based on City of Toronto's new 158 neighbourhood structure. See 'Table 5' for a sample of dataset.
```{r}
#| label: hc-by-nbh
#| fig-cap: "Table 5: Sample of Count of Hate Crime by Neighbourhood"
#| echo: false
#| 
# count each row by neighbourhood, and store the value in a new dataset
hc_by_nbh <- data %>% group_by(neighbourhood_158) %>%
  summarize(count = n())
# use 'knitr' generate a 10-row sample of the dataset
knitr::kable(head(hc_by_nbh, 10),
             col.names = c('Neighbourhood', 'Total Crimes'))
```


# Results {#sec-results}

The number of hate crime incidents has increased almost every year except 2019 and 2022. Last year's hate crimes surged to 329, nearly three times as much as the amount in 2018. The scatterplot intuitively shows a rapid growth rate(See 'Figure 1'). From the overall trend, the number of hate crimes is increasing every year.

```{r}
#| label: hc-year-scatter
#| fig-cap: "Number of Hate Crimes every year"
#| echo: false
#Use 'ggplot()' to generate the scatterplot based on the hc_by_year data
ggplot(hc_by_year, aes(x = occurrence_year, y =count)) +
  # Add scatterplot points
  geom_point(color='darkred') +
  # Add line connecting the points
  geom_line(color='darkred') +  
  # Add axis labels and title
  labs(
       x = "Year",
       y = "Hate Crimes",
       ) +
  theme_minimal()
```

The average of hate crime by each bias type is 193.5 and the median is 142.5. We have showed that there is no outlier based on the interquantile range test. From the graph (see Figure 2) we can say most of the causes are ethnicity, sex, race and religion. More particularly, religion bias is the number one cause for hate crime, and is way more than other bias.

```{r}
#| label: hc-bias-bar
#| fig-cap: "Hate Crime by Bias Type"
#| echo: false
ggplot(hc_by_bias, aes(x = reorder(bias_type,count), y = count)) +
#Create bar plot, choose the color
  geom_col(fill = "darkred", color = "black") + 
#Add name for x,y axis
  labs(x = "Bias Type", 
       y = "Count") +
  theme_minimal() 
```

The distribution by neighbourhood tells more stories. Since there are 158 neighbourhoods in Toronto, there are about 158 groups if I want to group by variable 'neighbourhood_158', which is too messy for a bar plot. So I use histogram to show the relationship(See Figure 3). The histogram is right-skewed and unimodal.From the graph we can find there are 69 neighbourhood that are have hate crime less than or equal to 5 from 2018 to 2023.There are 6 neighbourhoods have more than 25 crimes.

```{r}
#| label: hc-neighbourhood-bar
#| fig-cap: "Histogram of Hate Crime by Neighbourhood"
#| echo: false
ggplot(hc_by_nbh, aes(x = count)) +
  # Create a histogram. Set binwidth to 5 
  geom_histogram(binwidth = 5, fill = "darkred", color = "black", boundary = 0) +  
  # Set x-axis breaks to match bins
  scale_x_continuous(breaks = seq(0, 45, by = 5)) + 
  #Set y-axis breaks to match bins
  scale_y_continuous(breaks = seq(0, 80, by = 2))+ 
  #Give name for x,y axis
  labs(x = "Number of Hate Crimes", 
       y = "Number of Neighbourhoods") +
  theme_minimal()
```

# Discussion {#sec-disc}

These tables and figures give readers a visual and organized understanding of the development and causes of hate crimes. 'Figure 2' shows that most of the hate crimes are caused by religion bias, race bias and sex bias. 'Figure 3' shows that there are 6 neighbourhoods out of 158 are considered as dangerous neighbourhood while the rests have more than 25 hate crimes for recent 6 years. By revisiting the dataset 'hc_by_nbh' and sorting the dataset, I can find those 6 neighbourhoods are: Yonge-Bay Corridor, Church-Wellesley, Annex, Downtown Yonge East, York University Heights and Moss Park.

Algorithm-driven content recommendation systems often exacerbate polarization, leading users into place where different perspectives are diminished, and hate speech proliferates. This growing extremism can incite individuals or groups to escalate from online rhetoric to offline violence. As social media platforms continue to globalize and expand their user base, hate crimes may become more frequent and harder to control, driven by the influence of these platforms. In the near future, increasing in hate crime in Toronto is inevitable as more and more people of different nationalities and beliefs flock to this city, which is also showed in 'Figure 1'. So, the government should try its best to prevent hate crime.

The analyzed data and figures can help the government of Toronto understand the composition of hate crimes and the main areas where they occur. For example, the police can increase the number of patrols in dangerous areas and prevent crimes in time. The propaganda department can support views such as racial equality and religious freedom on social media and educate citizens to respect everyone.

# Conclusion{#sec-concl}

This paper investigated the hate crimes happened in Toronto from 2018 to 2023. The paper discussed the hate crimes' distribution in three aspects: year of occurrence, reason of occurrence, place of occurrence. It demonstrates that, the number of hate crimes is increasing in recent years; the top-three reason for hate crimes are religion bias, race bias and sex bias; among the 158 neighbourhoods in Toronto, 69 of them had less than 5 hate crimes in last 6 years and 6 of them had more than 25 hate crimes in the same period. In future study, I can analyze each bias deeper e.g. which religion experienced most hate crime so that the research can be more detailed.

# LLMs

Statement on LLM usage: no LLMs were used in the making of this paper.



\newpage


# References


