---
title: "Hate Crime in Toronto"
subtitle: "What is the Primary Reason for Hate Crime?"
author: 
  - Jingtao Luo
thanks: "Code and data are available at: LINK."
date: today
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# install.packages("tidyverse")
# install.packages("ggplot2")
# install.packages("dplyr")
# install.packages("readr")
# install.packages("janitor")
# install.packages("opendatatoronto")
# install.packages("knitr")
# install.packages("here")
library(tidyverse)
library(ggplot2)
library(dplyr)
library(readr)
library(janitor)
library(opendatatoronto)
library(knitr)
library(here)

#read the cleaned hate crime data
data <- read_csv(here::here("data/analysis_data/cleanned_data.csv"))


```


# Introduction

Toronto, often celebrated as one of the most multicultural cities in the world, stands as a beacon of diversity and inclusion. With over half of its population identifying as a visible minority and more than 180 languages spoken, the city reflects a rich tapestry of cultural, ethnic, and religious backgrounds. It has long been regarded as a model for successful urban integration, where immigrants and marginalized communities contribute to the city’s vibrant social, economic, and cultural fabric. However, beneath this inclusive narrative, Toronto also faces challenges in addressing the rising threat of hate crimes.

According to government of Canada [@govCan], hate crime is defined as criminal acts done by a person who is motivated by an extreme bias or hatred towards a particular social group. Research studies show that hate crimes cause “disproportionate harm” to individual victims as well as other members of the community belonging to the targeted social group. These crimes lead to not only physical but also psycho-emotional hurt. For example, if the assault occurs because the victim's race, then victim may live in fear of being victimized again for the rest of his life because a person's race can never be changed. 

Despite Toronto’s inclusive policies and progressive reputation, the city has not been immune to the global rise in hate-fueled violence. In recent years, incidents targeting specific communities—including Muslim, Jewish, Black, and LGBTQ+ populations—have underscored the persistent challenges that marginalized groups face. While hate crimes constitute only a fraction of overall crime, their impacts are disproportionately harmful, not only to direct victims but also to the broader communities they belong to, fostering fear, division, and alienation.

This paper aims to analyze the hate crimes in the following perspectives: when, where and why. When means the amount of crimes happened in each year. Where means the neighborhood of the hate crime occoured. And why is the reason that triggered this hate crime.

STILL MISSING!

# Data {#sec-data}

All relevant data can be found in the Open Data Toronto portal, and extracted using the `opendatatoronto` library for `R` [@rOpendatatoronto]. In particular, the dataset includes the 
information about each hate crime, investigated by the Hate Crime Unit by reported date since 2018.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false


```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it. 



# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.


### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```




# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows... 

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```



\newpage


# References


